---
title: "July Data"
output: html_document
editor_options: 
  chunk_output_type: console
---

## loading libraies
```{r}
library(tidyr)
library(ggplot2)
library(plyr)
library(MASS)
```
## Reading in the data and looking at its structure.
```{r}
data=read.csv("/Users/JamesB/Google Drive/Graduate School/Self_JamesTBrown/Data/Lipids/USDA BUP/20171102 July Samples RAW data.csv", header = T)
str(data)
```

## plotting out relevant data
there is a problem because the data is in column format. Especially for ggplot so we need to convert the trait data into columns and make it into 2 columns. 1 with trait ID and the other with trait measurements. this aloows us to call on the trait id column to plot.

Goal: plot larval data against season
```{r}
#converting witde to long
dat.long<-gather(data,traits,measurement,WET.Larva:Lipid.Wt)

## here are the plots
## all the traits against colony info and season as boxplot

ggplot(dat.long,aes(x=Season,y=measurement,colour=Colony.Info))+geom_boxplot()+facet_grid(traits~.,scales="free")

## all traits against cohort color coded by block with points
ggplot(dat.long,aes(x=Cohort,y=measurement,colour=factor(Block)))+geom_point()+facet_grid(traits~.,scales="free")

#subsetting out extreme sample lines in wet larvae 
subset(data,WET.Larva>0.6)

## removing data lines via subset
sub.out=subset(data,WET.Larva<0.6)
## or 
sub.out1=data[-95,]
dim(sub.out)
````

##Exploring the stats using ANOVA
````{r}
###making blocks as a factor
data$Block=as.factor(data$Block)

##constructing an aov model
mod1=aov(Lipid.Wt~Season*Colony.Info+Block+LEAN.Larvae,data=data)
summary(mod1)
`````
##Exploring the Stats using AIC
`````{r}
## checking model visualy
plot(mod1)

mod2=stepAIC(mod1, direction="both")
summary(mod2)
````
###Calculating the means. Visualized using boxplots and point plots 
````{r}
mean.dat=ddply(data,.(Season,Colony.Info),summarize,mean=mean(Lipid.Wt))
ggplot(data,aes(x=Season,y=Lipid.Wt,colour=Colony.Info))+geom_boxplot()+geom_line(data=mean.dat,aes(x=as.numeric(Season),y=mean,colour=Colony.Info))

ggplot(data,aes(x=Season,y=Lipid.Wt,colour=Colony.Info))+geom_point()+geom_line(data=mean.dat,aes(x=as.numeric(Season),y=mean,colour=Colony.Info))

data$resid<-scale(resid(aov(Lipid.Wt~Block+LEAN.Larvae,data=data)))
mean.dat2<-ddply(data,.(Season,Colony.Info),summarize,mean=mean(resid),sd=sd(resid))

#ggplot(data,aes(x=Season,y=resid,colour=Colony.Info))+geom_boxplot()#+geom_line(data=mean.dat,aes(x=as.numeric(Season),y=mean,colour=Colony.Info))

ggplot(mean.dat2,aes(x=Season,y=mean,colour=Colony.Info,group=Colony.Info))+geom_line(position="identity",size=2)+geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),width=.35,size=2)+geom_point(size=10)+ylab("Standardized Residuals of Lipid Weight")
`````

