---
title: "ECB Dry Pass Phenotype"
editor_options: null
  chunk_output_type: inline
chunk_output_type: inline
output: html_document
---
### Goal: Rear UZ and BE at 23C at 16:8 and 12:12 L:D regimens (UZ12, UZ16, BE12, BE16). Track dry mass changes during the 5th instar, the wandering stage, and during diapause. 

##Call Data
````{r}
library(readxl)
library(ggplot2)
library(nlme)
library(lme4)
library(lmerTest)
library(ez)
library(lsr)
library(dplyr)
library(tidyr)
library(MASS)
library(doBy)
library(plyr)


##PC
#data=read_excel("/Users/jbrown/Documents/GitHub/Lipid_Quantification/Data/Extraction Samples JLM.xlsx",sheet = "CONTSAMPdf")


##MAC
## reading in the data and looking at its structure
setwd("/Users/JamesB/Google Drive/Graduate School/Self_JamesTBrown/GitHub/Lipid_Quantification/Scripts_Analysis")
data=read_excel("/Users/JamesB/Google Drive/Graduate School/Self_JamesTBrown/GitHub/Lipid_Quantification/Data/Extraction Samples JLM.xlsx", sheet = "CONTSAMPdf")

data=subset(data,treat!="NA")
data=subset(data,cohort!="20180131")
data=subset(data,cohort!="20180206")
data=subset(data,sample_day!="W14")
data=subset(data,sample_day!="W19")
data=subset(data,sample_day!="W29")


##This code breaks apart the data in the column "treat" into "strain" and "photoperiod"
data$sample_id<-as.factor(as.character(data$sample_id))
data$strain<-substr(data$treat,1,2)
data$photoperiod<-as.numeric(substr(data$treat,3,4))
data$strain<-as.factor(as.character(data$strain))
data$photoperiod<-as.factor(as.character(data$photoperiod))
data$treat<-as.factor(as.character(data$treat))
data$sample_day<-factor(data$sample_day, levels=c("1" , "3" , "7" , "W" , "pW1" , "pW5" , "pW10" , "W14" ,"W15" , "pW18" , "W19" , "W20" , "W29" , "W30" , "pW38" , "W45"))
data$cohort<-as.factor(as.character(data$cohort))

#Reducing df
data1=data[1:626, c("sample_day","rep", "treat","sample_id","5th_date","cohort", "wet_mass","lean_mass","dry_mass","lipid_mass","strain","photoperiod")]

#converting wide to long. This code puts all the observations into one column with the label of your choice (measurements used below)
dataT=gather(data,traits,measurement,wet_mass:lipid_mass)
#str(dataT) #shows the structure of the data
#str(data)
````

##Traits Plots
````{r}
#Larva Trait Plot
dataTr=dataT
dataTr=subset(dataTr,measurement!="NA")
dataTr=subset(dataTr,photoperiod!="NA")
count(dataTr$treat)
dataTr=subset(dataTr,sample_day=="1"| sample_day=="W"| sample_day=="W15"| sample_day=="W20" | sample_day=="W30")
dataTr=subset(dataTr,measurement>=0)
dataTr=subset(dataTr,measurement<=0.1000)
ggplot(data=dataTr,aes(x=sample_day,y=measurement,fill=traits))+
  geom_boxplot()+
  facet_wrap(~traits)+
  theme_classic()+
  ggtitle("Larvae Traits")+
  ylab("Traits")+xlab("Sample Day")

#Larva lean Mass Plot
datal=data1
datal=subset(datal,lean_mass>=0)
datal=subset(datal,lean_mass<=0.1000)
count(datal$treat)
datal=subset(datal,sample_day=="1"| sample_day=="W"| sample_day=="W15"| sample_day=="W20" | sample_day=="W30")
```
##Dry Mass Comparison Between Treatments
```{r}
#Larva Dry Dry Plot
dataDrU=data1
#dataDrU=subset(dataDrU,treat=="UZ12"|treat=="UZ16")
dataDrU=subset(dataDrU,dry_mass>=0.0001)
dataDrU=subset(dataDrU,dry_mass<=0.1000)
count(dataDrU$treat)
dataDrU=subset(dataDrU,sample_day=="1" | sample_day=="W")
ggplot(data=dataDrU,aes(x=sample_day,y=dry_mass,color=treat,group=treat))+
  stat_summary(fun.y=mean, geom="line")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  theme_classic()+
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  ylab("Dry Mass")+xlab("")

#Larva Dry Mass Comparison Between Treatments
dataDry=data1
dataDry=subset(dataDry,dry_mass>=0)
#dataDry=subset(dataDry,dry_mass<=0.1000)
count(dataDry$treat)
dataDry=subset(dataDry,sample_day=="1"| sample_day=="W")

ggplot(data=dataDry,aes(x=sample_day,y=dry_mass,group=photoperiod,color=photoperiod))+
  stat_summary(aes(y = dry_mass), fun.y=mean, geom="line")+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 178 : BE 164 larvae)")+
  ylab("dry Mass")+xlab("Sample Day")




### Is there a significant difference in dry mass accumulation between treatments in preparation for diapause?
#### These Data include Day 1 larvae and Day W larvae from both treatments.

######################
# Larva Dry Mass    #
# Between Strains ####
# Between Treatments #
###### Day 1 #########


dataL1=subset(data,sample_day=="1")
dataL1=subset(dataL1,dry_mass>="0")
dataL1=subset(dataL1,dry_mass<="0.07")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Photoperiod Does NOT SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedL1=lmer(dry_mass ~ photoperiod + (1|rep) ,data=dataL1, REML = TRUE)
summary(mixedL1)
plot(mixedL1)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodL1=lmer(dry_mass ~ (1|rep),data=dataL1, REML = TRUE)
summary(submodL1)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataL1$stdresL1<-as.vector(scale(resid(submodL1)))

ggplot(data=dataL1,aes(x=sample_day,y=dry_mass,color=photoperiod))+
  stat_summary(aes(y = dry_mass), fun.y=mean)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  facet_wrap(~photoperiod)+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day 1")



######################
# Larva Dry Mass    #
# Between Strains ####
# Between Treatments #
###### Day W #########

##Mixed model with 'rep' as a random factor

dataLW=subset(data,sample_day=="W")
dataLW=subset(dataLW,dry_mass>="0")
dataLW=subset(dataLW,dry_mass<="0.07")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Photoperiod does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLW=lmer(dry_mass ~ photoperiod + (1|rep) ,data=dataLW, REML = TRUE)
summary(mixedLW)
plot(mixedLW)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLW=lmer(dry_mass ~ (1|rep),data=dataLW, REML = TRUE)
summary(submodLW)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataLW$stdresLW<-as.vector(scale(resid(submodLW)))

ggplot(data=dataLW,aes(x=sample_day,y=dry_mass,color=photoperiod))+
  stat_summary(aes(y = dry_mass), fun.y=mean)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  facet_wrap(~photoperiod)+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day W")
```

##Dry Mass Comparison Within Diapause Treatment and Between UZ12 and BE12 
```{r}
######################
#Larva dry Mass Plot#
# Within  Treatments #
######################
 
### Diapause###
dataDrD=data1
dataDrD=subset(dataDrD,photoperiod==12)
dataDrD=subset(dataDrD,dry_mass>=0)
#dataDrD=subset(dataDrD,dry_mass<=0.1000)
count(dataDrD$treat)
dataDrD=subset(dataDrD,sample_day=="1"| sample_day=="W")

ggplot(data=dataDrD,aes(x=sample_day,y=dry_mass,group=treat,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Diapause: Between Strain Comparison of Larvae dry Mass (UZ 117 : BE 107 larvae)")+
  ylab("dry Mass")+xlab("Sample Day")

dataDrDl=data1
dataDrDl=subset(dataDrDl,photoperiod==12)
dataDrDl=subset(dataDrDl,dry_mass>=0)
#dataDrDl=subset(dataDrDl,dry_mass<=0.1000)
count(dataDrDl$treat)
dataDrDl=subset(dataDrDl,sample_day=="W"| sample_day=="W15"| sample_day=="W20" | sample_day=="W30")

ggplot(data=dataDrDl,aes(x=sample_day,y=dry_mass,group=treat,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Diapause: Between Strain Comparison of Larvae dry Mass (UZ 117 : BE 107 larvae)")+
  ylab("dry Mass")+xlab("Sample Day")

######################
# Larva Dry Mass    #
# Within  Treatments #
###### Day 1 #########

##Mixed model with 'rep' as a random factor

dataDrD1=subset(data1,photoperiod==12)
dataDrD1=subset(dataDrD1,dry_mass>=0)
dataDrD1=subset(dataDrD1,sample_day=="1")
count(dataDrD1$treat)

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeD1=lmer(dry_mass ~  strain + (1|rep),data=dataDrD1, REML = TRUE)
summary(mixedLeD1)
plot(mixedLeD1)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeD1=lmer(dry_mass ~ (1|rep),data=dataDrD1, REML = TRUE)
summary(submodLeD1)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrD1$stdresLeD1<-as.vector(scale(resid(submodLeD1)))

ggplot(data=dataDrD1,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 25 : BE 17 larvae)")+
  ylab("dry Mass")+xlab("Sample Day 1")

######################
# Larva Dry Mass    #
# Between Strains ####
# Within  Treatments #
###### Day W #########

##Mixed model with 'rep' as a random factor

dataDrDW=subset(data1,photoperiod==12)
dataDrDW=subset(dataDrDW,dry_mass>=0)
dataDrDW=subset(dataDrDW,sample_day=="W")


#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeDW=lmer(dry_mass ~  strain + (1|rep),data=dataDrDW, REML = TRUE)
summary(mixedLeDW)
plot(mixedLeDW)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeDW=lmer(dry_mass ~ (1|rep),data=dataDrDW, REML = TRUE)
summary(submodLeDW)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrDW$stdresLeDW<-as.vector(scale(resid(submodLeDW)))

ggplot(data=dataDrDW,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 34 : BE 43 larvae)")+
  ylab("dry Mass")+xlab("Sample Day W")


######################
# Larva Dry Mass    #
# Between Strains ####
# Within  Treatments #
###### Day W15 #######

##Mixed model with 'rep' as a random factor

dataDrDW15=subset(data1,sample_day=="W15")
dataDrDW15=subset(dataDrDW15,dry_mass>="0")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Photoperiod does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeDW15=lmer(dry_mass ~ strain + (1|rep) ,data=dataDrDW15, REML = TRUE)
summary(mixedLeDW15)
plot(mixedLeDW15)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeDW15=lmer(dry_mass ~ (1|rep),data=dataDrDW15, REML = TRUE)
summary(submodLeDW15)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrDW15$stdresLeDW15<-as.vector(scale(resid(submodLeDW15)))

ggplot(data=dataDrDW15,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 19 : BE 37 larvae)")+
  ylab("dry Mass")+xlab("Sample Day W15")

######################
# Larva Dry Mass    #
# Between Strains ####
# Within  Treatments #
###### Day w20 #######

##Mixed model with 'rep' as a random factor

dataDrDW20=subset(data1,sample_day=="W20")
dataDrDW20=subset(dataDrDW20,dry_mass>="0")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Photoperiod does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeDW20=lmer(dry_mass ~ strain + (1|rep) ,data=dataDrDW20, REML = TRUE)
summary(mixedLeDW20)
plot(mixedLeDW20)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeDW20=lmer(dry_mass ~ (1|rep),data=dataDrDW20, REML = TRUE)
summary(submodLeDW20)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrDW20$stdresLeDW20<-as.vector(scale(resid(submodLeDW20)))

ggplot(data=dataDrDW20,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 20 : BE 6 larvae)")+
  ylab("dry Mass")+xlab("Sample Day W20")


######################
# Larva Dry Mass    #
# Between Strains ####
# Within  Treatments #
###### Day w30 #######

##Mixed model with 'rep' as a random factor

dataDrDW30=subset(data1,sample_day=="W30")
dataDrDW30=subset(dataDrDW30,dry_mass>="0")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Photoperiod does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeDW30=lmer(dry_mass ~ strain + (1|rep) ,data=dataDrDW30, REML = TRUE)
summary(mixedLeDW30)
plot(mixedLeDW30)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeDW30=lmer(dry_mass ~ (1|rep),data=dataDrDW30, REML = TRUE)
summary(submodLeDW30)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrDW30$stdresLeDW30<-as.vector(scale(resid(submodLeDW30)))

ggplot(data=dataDrDW30,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 19 : BE 4 larvae)")+
  ylab("dry Mass")+xlab("Sample Day W30")
```

##Dry Mass Comparison Within Non Diapause Treatment and Between UZ16 and BE16 
```{r}
######################
#Larva dry Mass Plot#
# Within   Treatment #
######################

### Non Diapause###
dataDrND=data1
dataDrND=subset(dataDrND,photoperiod==16)
dataDrND=subset(dataDrND,dry_mass>=0)
#dataDrND=subset(dataDrND,dry_mass<=0.1000)
count(dataDrND$treat)
dataDrND=subset(dataDrND,sample_day=="1"| sample_day=="W")

ggplot(data=dataDrND,aes(x=sample_day,y=dry_mass,group=treat,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Non Diapause: Between Strain Comparison of Larvae dry Mass (UZ 117 : BE 107 larvae)")+
  ylab("dry Mass")+xlab("Sample Day")
  
######################
# Larva Dry Mass    #
# Within  Treatments #
###### Day 1 #########

##Mixed model with 'rep' as a random factor

dataDrND1=subset(data1,photoperiod==16)
dataDrND1=subset(dataDrND1,dry_mass>=0)
dataDrND1=subset(dataDrND1,sample_day=="1")
count(dataDrND1$treat)

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeND1=lmer(dry_mass ~  strain + (1|rep),data=dataDrND1, REML = TRUE)
summary(mixedLeND1)
plot(mixedLeND1)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeND1=lmer(dry_mass ~ (1|rep),data=dataDrND1, REML = TRUE)
summary(submodLeND1)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrND1$stdresLeND1<-as.vector(scale(resid(submodLeND1)))

ggplot(data=dataDrND1,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 25 : BE 17 larvae)")+
  ylab("dry Mass")+xlab("Sample Day 1")

######################
# Larva Dry Mass    #
# Between Strains ####
# Within  Treatments #
###### Day W #########

##Mixed model with 'rep' as a random factor

dataDrNDW=subset(data1,photoperiod==12)
dataDrNDW=subset(dataDrNDW,dry_mass>=0)
dataDrNDW=subset(dataDrNDW,sample_day=="W")


#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeNDW=lmer(dry_mass ~  strain + (1|rep),data=dataDrNDW, REML = TRUE)
summary(mixedLeNDW)
plot(mixedLeNDW)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeNDW=lmer(dry_mass ~ (1|rep),data=dataDrNDW, REML = TRUE)
summary(submodLeNDW)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrNDW$stdresLeNDW<-as.vector(scale(resid(submodLeNDW)))

ggplot(data=dataDrNDW,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("Treatment Comparison of Larvae dry Mass (UZ 34 : BE 43 larvae)")+
  ylab("dry Mass")+xlab("Sample Day W")
```

##Dry Mass Comparison Between treatments Within BE strain
```{r}
######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
######################

### BE ###
dataDrBE=data1
dataDrBE=subset(dataDrBE,strain=="BE")
dataDrBE=subset(dataDrBE,dry_mass>=0)
#dataDrBE=subset(dataDrBE,dry_mass<=0.1000)
count(dataDrBE$treat)
dataDrBE=subset(dataDrBE,sample_day=="1"| sample_day=="W")

ggplot(data=dataDrBE,aes(x=sample_day,y=dry_mass,group=treat,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("BE: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day")
  

######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
###### Day 1 #########

##Mixed model with 'rep' as a random factor

dataDrBED1=data1
dataDrBED1=subset(dataDrBED1,strain=="BE")
dataDrBED1=subset(dataDrBED1,dry_mass>=0)
dataDrBED1=subset(dataDrBED1,sample_day=="1")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeBED1=lmer(dry_mass ~  treat + (1|rep),data=dataDrBED1, REML = TRUE)
summary(mixedLeBED1)
plot(mixedLeBED1)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeBED1=lmer(dry_mass ~ (1|rep),data=dataDrBED1, REML = TRUE)
summary(submodLeBED1)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrBED1$stdresLeBED1<-as.vector(scale(resid(submodLeBED1)))

ggplot(data=dataDrBED1,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("BE: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day 1")

######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
###### Day W #########

##Mixed model with 'rep' as a random factor

dataDrBEDW=data1
dataDrBEDW=subset(dataDrBEDW,strain=="BE")
dataDrBEDW=subset(dataDrBEDW,dry_mass>=0)
dataDrBEDW=subset(dataDrBEDW,sample_day=="W")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeBEDW=lmer(dry_mass ~  treat + (1|rep),data=dataDrBEDW, REML = TRUE)
summary(mixedLeBEDW)
plot(mixedLeBEDW)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeBEDW=lmer(dry_mass ~ (1|rep),data=dataDrBEDW, REML = TRUE)
summary(submodLeBEDW)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrBEDW$stdresLeBEDW<-as.vector(scale(resid(submodLeBEDW)))

ggplot(data=dataDrBEDW,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",BE12="orange",BE16="orange"))+
  theme_classic()+
  ggtitle("BE: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day W")
```
##Dry Mass Comparison Between treatments Within UZ strain
```{r}
######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
######################

### UZ ###
dataDrUZ=data1
dataDrUZ=subset(dataDrUZ,strain=="UZ")
dataDrUZ=subset(dataDrUZ,dry_mass>=0)
#dataDrUZ=subset(dataDrUZ,dry_mass<=0.1000)
count(dataDrUZ$treat)
dataDrUZ=subset(dataDrUZ,sample_day=="1"| sample_day=="W")

ggplot(data=dataDrUZ,aes(x=sample_day,y=dry_mass,group=treat,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",UZ12="orange",UZ16="orange"))+
  theme_classic()+
  ggtitle("UZ: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day")
  

######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
###### Day 1 #########

##Mixed model with 'rep' as a random factor

dataDrUZD1=data1
dataDrUZD1=subset(dataDrUZD1,strain=="UZ")
dataDrUZD1=subset(dataDrUZD1,dry_mass>=0)
dataDrUZD1=subset(dataDrUZD1,sample_day=="1")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeUZD1=lmer(dry_mass ~  treat + (1|rep),data=dataDrUZD1, REML = TRUE)
summary(mixedLeUZD1)
plot(mixedLeUZD1)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeUZD1=lmer(dry_mass ~ (1|rep),data=dataDrUZD1, REML = TRUE)
summary(submodLeUZD1)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrUZD1$stdresLeUZD1<-as.vector(scale(resid(submodLeUZD1)))

ggplot(data=dataDrUZD1,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",UZ12="orange",UZ16="orange"))+
  theme_classic()+
  ggtitle("UZ: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day 1")

######################
# Larva dry Mass    #
# Within Strain      #
# Between Treatments #
###### Day W #########

##Mixed model with 'rep' as a random factor

dataDrUZDW=data1
dataDrUZDW=subset(dataDrUZDW,strain=="UZ")
dataDrUZDW=subset(dataDrUZDW,dry_mass>=0)
dataDrUZDW=subset(dataDrUZDW,sample_day=="W")

#this model attempts to fit the linear mixed effects model with 3 variables; photoperiod, dry mass, and the random effect of rep
## Strain does SIGNIFICANTLY effect dry mass accumulation between larvae reared in long day vs short day conditions
mixedLeUZDW=lmer(dry_mass ~  treat + (1|rep),data=dataDrUZDW, REML = TRUE)
summary(mixedLeUZDW)
plot(mixedLeUZDW)
##The following model removes photoperiod as factors to determine the effect of dry mass and Rep on dry mass
submodLeUZDW=lmer(dry_mass ~ (1|rep),data=dataDrUZDW, REML = TRUE)
summary(submodLeUZDW)

## Standard Residuals -  Adding stdres to df and checking for regularity of distribution within +/-2 standard deviations  
dataDrUZDW$stdresLeUZDW<-as.vector(scale(resid(submodLeUZDW)))

ggplot(data=dataDrUZDW,aes(x=sample_day,y=dry_mass,color=treat))+
  stat_summary(fun.y=mean, geom="line",shape=18)+
  stat_summary(fun.data=mean_se, geom="errorbar", width=0.05) +
  scale_color_manual(values=c(UZ12="purple",UZ16="purple",UZ12="orange",UZ16="orange"))+
  theme_classic()+
  ggtitle("UZ: Within Strain Comparison of Larvae dry Mass")+
  ylab("dry Mass")+xlab("Sample Day W")
```